---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r loadpreprocessing, echo = T}
dir <- "D:/Online courses/Coursera courses/Data Science Specialization/05- Reproducible research/Assignment_1/RepData_PeerAssessment1"
setwd(dir)
# Load the data after unzip it
data <- read.csv(unz(paste(dir,"/activity.zip",sep = ""),filename = "activity.csv"))
require(dplyr,quietly = T, warn.conflicts = FALSE)
# Sum the number of steps per day to plot the histogram
sumperday <- data %>% group_by(date) %>% summarise_each(funs(sum))

```


## What is mean total number of steps taken per day?
```{r histogram1, echo = T}
# Plot the histogram
hist(sumperday$steps, xlab = "Number of steps", main = "Total number of steps taken each day")
```

The following table shows the mean and median of the number of steps taken each day.

```{r showtable, echo = T, results = "asis"}
# Calculate the mean and median of the number of steps taken each day
meanandmeadian <- data %>% group_by(date) %>% summarise_each_(funs(mean,median),"steps")

require(xtable,quietly = T)
xt <- xtable(meanandmeadian)
print(xt, include.rownames=FALSE,type = "html")
```


## What is the average daily activity pattern?
```{r timeseriesplot, echo = T}
# Calculate the average of steps taken in the 5 min interval across days
meaninterval <- data %>% group_by(interval) %>% summarise_each_(funs(mean(., na.rm = T)),"steps")
#plot the time series
with(meaninterval,plot(interval,steps, type = "l",xlab = "Interval", ylab= "Number of steps", main = "Average number of steps taken in each 5 min interval across all days"))
# Find the interval at which the number of steps is the maximum across all days
maxint <- meaninterval[with(meaninterval,which.max(steps)),1]
```

The `r maxint`<sup>th</sup> interval is the on in which the number of steps is maximum across all days.


## Imputing missing values

```{r NAnumber, echo = T}
# Find out the number of NA in the data
nanumber <- length(with(data,which(is.na(steps)== T)))
```

The number of non available (NA) data along all days and intervals is `r nanumber`.

To replace the missing values, the data was grouped by each 5 min interval across days. The NA values were replaced by the mean of these 5 min intervals.

```{r replaceNA, echo = T}
# Create a new dataframe replacing all the NAs with the mean by interval
newdata <- data %>% group_by(interval) %>% mutate(steps = replace(steps,is.na(steps),mean(steps,na.rm = T)))
```

The following figure shows the re-made histogram with all the missing values replaced by the mean of each interval across all days.


```{r histogram2, echo = T}
# Sum the number of steps per day to plot the histogram
sumperday2 <- newdata %>% group_by(date) %>% summarise_each(funs(sum))
# Plot histogram with new dataset
hist(sumperday2$steps, xlab = "Number of steps", main = "Total number of steps taken each day")
```

The following table shows the mean and median of the number of steps taken each day after replacing the NA values with average values.


```{r showtable2, echo = T, results = "asis"}
# Calculate the mean and median of the number of steps taken each day
meanandmeadian2 <- newdata %>% group_by(date) %>% summarise_each_(funs(mean,median),"steps")

require(xtable,quietly = T)
xt2 <- xtable(meanandmeadian2)
print(xt2, include.rownames=FALSE,type = "html")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays, echo = T}
# Create the new factor variable with two levels = Weekday and Weekend, the calculate the average number of steps in each interval across weedays and weekend
weekdaysdf <- newdata %>% mutate(weekday = factor(ifelse(weekdays(as.Date(date)) == "Saturday" | weekdays(as.Date(date)) =="Sunday","Weekend","Weekday"))) %>% group_by(weekday,interval) %>% summarise_each_(funs(mean),"steps")

# Create the plot showing the average number of steps in each interval for the weekends and weekdays
require(ggplot2, quietly = T)
ggplot(weekdaysdf,aes(interval,steps)) + geom_line() + facet_grid(weekday~.) + ggtitle("Average number of steps taken in each 5 min interval across \n weekdays and weekends")

```
